---
title: "Final Project 401"
author: "Stone Cai"
date: "2024-10-20"
output:
  pdf_document: default
  html_document: default
---

```{r}

data<- read.csv ("C:/Users/stone/Documents/Stat 401/Final Project data/LeagueofLegends.csv")

##View(data)
head(data)

```



First I want to filter to only keep variables relavent to my analysis of early game advantages. 
```{r}
library(dplyr)
df_subset <- data%>% select(Year, bResult, rResult,golddiff,bKills,bTowers,bDragons,bHeralds,rKills, rTowers,rDragons,rHeralds)

View(df_subset)

```

On thing I notice is that all 2015 games do not have any entries for the objective called "Herald" was introduced into the game in 2016. Thus, if I want to include herald as a
final variable in my regression, I have to exclude all games from 2015. 

```{r}
df_filtered <- df_subset %>% filter(Year != 2015)
##View(df_filtered)

```
Next, most of my variables have elements that are actually lists. I am going to extract the gold diff at 15 minutes. For variables such as kills, towers or dragons, I want to extract the first instance blue team secures the objective and the first instance red.


```{r}
df_filtered1 <- df_filtered %>%
  mutate(golddiff1 = NA, bKills1=NA,bTowers1=NA, bDragons1=NA, bHeralds1=NA, rKills1=NA,rTowers1=NA, rDragons1=NA, rHeralds1=NA,)

## extract gold difference at 15 minutes



df_filtered1$golddiff1 <- sapply(df_filtered1$golddiff, function(golddiff_string) {
  #golddiff_string <- df_filtered1$golddiff[[1]]
  cleaned_string <- gsub("\\[|\\]", "", golddiff_string)
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]] 
  numeric_values <- as.numeric(split_values)
 
  
   return(numeric_values[15])
})

View(df_filtered1)

##dont take absolute value of gold lead. If positive blue team is ahead at 15 mintues if negative blue team is behind at 15 minutes. 


View(df_filtered1)


##########################



df_filtered1$bKills1 <- sapply(df_filtered1$bKills, function(bluekills_string) {
  
  #bkills_string <- df_filtered1$bKills[[1]]
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", bluekills_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)

### cleaning blueTowers

df_filtered1$bTowers1 <- sapply(df_filtered1$bTowers, function(bluetowers_string) {
  
  #bkills_string <- df_filtered1$bKills[[1]]
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", bluetowers_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})

##View(df_filtered1)

#####

##View(df_filtered1)

### cleaning blueTowers

df_filtered1$bDragons1 <- sapply(df_filtered1$bDragons, function(bluedragons_string) {
  
  #bkills_string <- df_filtered1$bKills[[1]]
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", bluedragons_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
View(df_filtered1)


```