---
title: "Final Project 401"
author: "Stone Cai"
date: "2024-10-20"
output:
  pdf_document: default
  html_document: default
---

```{r}

data<- read.csv ("C:/Users/stone/Documents/Stat 401/Final Project data/LeagueofLegends.csv")

##View(data)
head(data)

```



First I want to filter to only keep variables relavent to my analysis of early game advantages. 
```{r}
library(dplyr)
df_subset <- data%>% select(Year, bResult, rResult,golddiff,bKills,bTowers,bDragons,bHeralds,rKills, rTowers,rDragons,rHeralds)

View(df_subset)

```

On thing I notice is that all 2015 games do not have any entries for the objective called "Herald" which 
was introduced into the game in 2016. Thus, if I want to include herald as a
final variable in my regression, I have to exclude all games from 2015. 

```{r}
df_filtered <- df_subset %>% filter(Year != 2015)
##View(df_filtered)

```
Next, most of my variables have elements that are actually lists. I am going to extract the gold diff at 15 minutes. For variables such as kills, towers or dragons, I want to extract the first instance blue team secures the objective and the first instance red.


```{r}
df_filtered1 <- df_filtered %>%
  mutate(golddiff1 = NA, bKills1=NA,bTowers1=NA, bDragons1=NA, bHeralds1=NA, rKills1=NA,rTowers1=NA, rDragons1=NA, rHeralds1=NA,)

## extract gold difference at 15 minutes



df_filtered1$golddiff1 <- sapply(df_filtered1$golddiff, function(golddiff_string) {
  #golddiff_string <- df_filtered1$golddiff[[1]]
  cleaned_string <- gsub("\\[|\\]", "", golddiff_string)
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]] 
  numeric_values <- as.numeric(split_values)
 
  
   return(numeric_values[15])
})


df_filtered1$golddiff1<- df_filtered1$golddiff1/ 1000

##View(df_filtered1)

##dont take absolute value of gold lead. If positive blue team is ahead at 15 mintues if negative blue team is behind at 15 minutes. 




##########################



df_filtered1$bKills1 <- sapply(df_filtered1$bKills, function(bluekills_string) {
  
  #bkills_string <- df_filtered1$bKills[[1]]
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", bluekills_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)

### cleaning blueTowers

df_filtered1$bTowers1 <- sapply(df_filtered1$bTowers, function(bluetowers_string) {
  
  #bkills_string <- df_filtered1$bKills[[1]]
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", bluetowers_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})

########

### cleaning blueTowers

df_filtered1$bDragons1 <- sapply(df_filtered1$bDragons, function(bluedragons_string) {
  
  #bkills_string <- df_filtered1$bKills[[1]]
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", bluedragons_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)

##clean blue heralds

df_filtered1$bHeralds1 <- sapply(df_filtered1$bHeralds, function(blueheralds_string) {
  
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", blueheralds_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)

##clean red kills

df_filtered1$rKills1 <- sapply(df_filtered1$rKills, function(redkills_string) {
  
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", redkills_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)

### clean red towers

df_filtered1$rTowers1 <- sapply(df_filtered1$rTowers, function(redtowers_string) {
  
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", redtowers_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)
###red dragons1

df_filtered1$rDragons1 <- sapply(df_filtered1$rDragons, function(reddragons_string) {
  
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", reddragons_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)

##red heralds
df_filtered1$rHeralds1 <- sapply(df_filtered1$rHeralds, function(redheralds_string) {
  
    # Check if the input is NULL or empty

  cleaned_string <- gsub("\\[|\\]", "", redheralds_string)
  
  if (cleaned_string == "") {
    return(NA)  # Return NA for invalid inputs
  }
  split_values <- strsplit(cleaned_string, ",\\s*")[[1]]

  
  numeric_values <- as.numeric(split_values[1])
  #as.numeric(split_values[1])
 
  
   return(numeric_values)
})
##View(df_filtered1)



```

Here,I clean each of my variables I'm interested in. For gold diff, I took bluegold diff at 15 minutes. So given a list of gold difference at every minute in a game, I chose the gold lead at 15 minutes as proxy
to predict odds of winning the game. Note: Every game sampled did last at least 15 minutes so this is a reasonable proxy. Next, I wanted to extract the time stamps of when first kill, tower, dragon, and herald occur. To do so, I need to get time stamps for both blue and red teams. These variables are all list of lists. The list is sorted by increasing time already so I just need to extract the first list split the strings within each list and extract the first ELEMENT of each list. I have done so above. 


```{r}
df_filtered2 <- df_filtered1 %>%
  mutate( Killdummy=NA,Towerdummy=NA, Dragondummy=NA, Heralddummy=NA)

##View(df_filtered2)


# Add result column
df_filtered2$Killdummy <- ifelse(
  is.na(df_filtered2$bKills1), 0,  # If bKills1 is NA, return 0
  ifelse(is.na(df_filtered2$rKills1), 1,  # If rKills1 is NA, return 1
         ifelse(df_filtered2$bKills1 < df_filtered2$rKills1, 1, 0)  # Otherwise, compare values
  )
)


# Towerdummy
df_filtered2$Towerdummy <- ifelse(
  is.na(df_filtered2$bTowers1), 0,  # If bTower is NA, return 0
  ifelse(is.na(df_filtered2$rTowers1), 1,  # If rTower is NA, return 1
         ifelse(df_filtered2$bTowers1 < df_filtered2$rTowers1, 1, 0)  # Otherwise, compare values
  )
)

# Dragondummy
df_filtered2$Dragondummy <- ifelse(
  is.na(df_filtered2$bDragons1), 0,  
  ifelse(is.na(df_filtered2$rDragons1), 1, 
         ifelse(df_filtered2$bDragons1 < df_filtered2$rDragons1, 1, 0)  # Otherwise, compare values
  )
)

length(df_filtered2$bDragons1)


# Heralddummy
df_filtered2$Heralddummy <- ifelse(
  is.na(df_filtered2$bHeralds1) & is.na(df_filtered2$rHeralds1), NA,  # Both are NA
  ifelse(is.na(df_filtered2$bHeralds1), 0,  # Only bHerald1 is NA
         ifelse(is.na(df_filtered2$rHeralds1), 1,  # Only rHerald1 is NA
                ifelse(df_filtered2$bHeralds1 < df_filtered2$rHeralds1, 1, 0)  # Compare values
         )
  )
)



##View (df_filtered2)

```


```{r}
df_filtered3 <- df_filtered2 %>%
  mutate( FBlood=NA,FTower=NA, FDragon=NA)




df_filtered3$FBlood <- ifelse(
  is.na(df_filtered3$bKills1), df_filtered3$rKills1,  # If bKills1 is NA, return 0
  ifelse(is.na(df_filtered3$rKills1),df_filtered3$ bKills1,  # If rKills1 is NA, return 1
         ifelse(df_filtered3$bKills1 < df_filtered2$rKills1, df_filtered3$ bKills1, df_filtered3$rKills1)  # Otherwise, compare values
  )
)


##repeat for first tower


df_filtered3$FTower <- ifelse(
  is.na(df_filtered3$bTowers1), df_filtered3$rTowers1,  # If bKills1 is NA, return 0
  ifelse(is.na(df_filtered3$rTowers1),df_filtered3$ bTowers1,  # If rKills1 is NA, return 1
         ifelse(df_filtered3$bTowers1 < df_filtered2$rTowers1, df_filtered3$ bTowers1, df_filtered3$rTowers1)  # Otherwise, compare values
  )
)


##repeat for dragon




df_filtered3$FDragon <- ifelse(
  is.na(df_filtered3$bDragons1), df_filtered3$rDragons1,  # If bKills1 is NA, return 0
  ifelse(is.na(df_filtered3$rDragons1),df_filtered3$ bDragons1,  # If rKills1 is NA, return 1
         ifelse(df_filtered3$bDragons1 < df_filtered2$rDragons1, df_filtered3$ bDragons1, df_filtered3$rDragons1)  # Otherwise, compare values
  )
)





View(df_filtered3)

```


